### 2022-06-23 analyze.Rmd

---
title: "ARTIS S-net Overview"
output: html_document
date: '2022-06-21'
author: "Jessica Gephart and Rahul Agrawal Bejarano"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Reset Environment to start with an empty workspace
rm(list=ls())

library(tidyverse)
library(countrycode)
library(kableExtra)

# POINT THESE paths to folders and files in your local computer
# File Path to the folder where your data lives
datadir <- "data"
# File Path to the folder where you want to ouput results
outdir <- "Outputs"

# File name of the data you want to use
artis_snet_filename <- "20220616_josephbarnes_ARTIS_snet.csv"

```

```{r include=FALSE}

artis <- read.csv(file.path(datadir, artis_snet_filename))
```

## Timeseries

### Creating a timeseries line graph

Below is a code sample for creating timeseries by year based on seafood product weights.
```{r }
# Function that creates a timeseries
# INPUT:
#   - df: A data frame with a Year column and Product weight
#   x_var and y_var need to be passed in tidyverse style as variables
#   - x_var: col variable that will be plotted on the x axis (Group by var)
#   - y_var: col variable that will be plotted on the y axis (Summarize var)
# OUTPUTS:
#   - p: Timeseries plot
create_timeseries_line <- function(df, x_var, y_var, plot_title, x_title, y_title) {
  timeseries <- df %>%
  group_by({{ x_var }}) %>%
  summarize(quantity = sum({{ y_var }}))

  p <- ggplot(timeseries) +
    geom_line(aes(x={{ x_var }}, y=quantity)) +
    labs(x=x_title, y=y_title, title=plot_title)

  return(list(p, timeseries))
}

timeseries_results <- artis %>% 
  create_timeseries_line(x_var=year, y_var=product_weight_t,
                              "Timeseries of ARTIS S-net HS12", 
                              "Year", "Quantity (tonnes)")

timeseries_p <- timeseries_results[[1]]
timeseries_df <- timeseries_results[[2]]

rm(timeseries_results)

plot(timeseries_p)

# Use the code below to get a visual representation of the timeseries as a dataframe
# timeseries_df %>%
#   kable() %>%
#   kable_styling(full_width=FALSE)

```

### Creating a stacked line graph based on Species, HS codes, Trade Partners
```{r}
# Creates a stacked graph based on a time variable and stacking variable
# INPUTS
#   - df: dataframe containing data to be visualized
#   - time_var: X axis variables
#   - stack_var: variable to create stacking groups
#   - q_var:Y axis variable
#   - prop_flow_cutoff: Minimum proportion cutoff, 
#       - all groups below this cutoff get grouped into an "Other" group
# OUTPUTS:
#   - p: plot of a stacked timeseries
create_stacked_graph <- function(df, time_var, stack_var, q_var, 
                                 prop_flow_cutoff=0.05, 
                                 plot_title="", x_title="", y_title="") {

  stacked <- df %>%
    group_by({{ time_var }}, {{ stack_var }}) %>%
    summarize(quantity = sum({{ q_var }})) %>%
    ungroup() %>%
    group_by({{ time_var }}) %>%
    mutate(total_by_time = sum(quantity)) %>%
    mutate(prop_flow = quantity / total_by_time) %>%
    mutate(common_name = case_when(
      prop_flow < prop_flow_cutoff ~ "Other",
      TRUE ~ as.character({{ stack_var }})
    )) %>%
    ungroup() %>%
    group_by({{ time_var }}, common_name) %>%
    summarize(quantity = sum(quantity)) %>%
    ungroup()

  year_grid <- expand_grid(common_name=unique(stacked$common_name),
                           year=unique(stacked %>% pull({{ time_var }})))

  stacked <- stacked %>%
    full_join(year_grid) %>%
    mutate(quantity = case_when(
      is.na(quantity) ~ 0,
      TRUE ~ as.numeric(quantity)
    ))

  p <- stacked %>%
    ggplot() +
    geom_area(aes(x={{ time_var }}, y=quantity, fill=common_name)) +
    scale_fill_viridis_d() +
    labs(x=x_title, y=y_title, title=plot_title, fill="Common Name") +
    theme_bw()

  return(list(p, stacked))
}
```

An example using species as the stacking group:
```{r message=FALSE}

species_stacked <- artis %>% 
  create_stacked_graph(year, SciName, product_weight_t, 
                            plot_title="Species breakdown over time", 
                            x_title="Year", y_title="Quantity (tonnes)")

species_p <- species_stacked[[1]]
species_df <- species_stacked[[2]]

rm(species_stacked)

plot(species_p)
```

An example using hs product codes as the stacking group:
```{r message=FALSE}

hs_stacked <- artis %>% 
       create_stacked_graph(year, hs6, product_weight_t, 
                            plot_title="Seafood product (by HS code) over time", 
                            x_title="Year", y_title="Quantity (tonnes)")

hs_p <- hs_stacked[[1]]
hs_df <- hs_stacked[[2]]

rm(hs_stacked)

plot(hs_p)
```

An example using exporter as the stacking group:
```{r message=FALSE}

exporter_stacked <- artis %>% 
       create_stacked_graph(year, exporter_iso3c, product_weight_t, 
                            plot_title="Exporters over time", 
                            x_title="Year", y_title="Quantity (tonnes)")

exporter_p <- exporter_stacked[[1]]
exporter_df <- exporter_stacked[[2]]

rm(exporter_stacked)

plot(exporter_p)
```