### 2024-03-14 ms-russia-figures.Rmd
change panel labels

---
title: "Russian sanctions figures"
output: html_document
date: "2024-02-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To do: 
-

```{r}
library(tidyverse)
library(countrycode)
library(exploreARTIS)
library(DBI)
library(janitor)
library(broom)
library(dbscan)
library(GGally)
library(ggrepel)
library(ggpubr)
```

```{r}
habitat_source_colors <- c("#1B708F", # blue = marine capture
                          "#86ADA7", # light blue = freshwater capture
                          "#F37542", # medium orange = marine aquaculture
                          "#F7AF75", # light orange = inland aquaculture
                          "grey50" # unknown
                          )

region_colors <- c(
  "#741A32", # maroon
  "#B34232", # burnt orange
  "#D38F35", #  orange
  "#D4B95f", # khaki
  "#4FA2A2", # teal
  "#114F59" # dark teal
)
```

```{r}
# Initial database pulls

# Database connection
con <- dbConnect(RPostgres::Postgres(),
                 dbname=Sys.getenv("DB_NAME"),
                 host="localhost",
                 port="5432",
                 user=Sys.getenv("DB_USERNAME"),
                 password=Sys.getenv("DB_PASSWORD"))

# Check that connection is established by checking which tables are present
dbListTables(con)

# ARTIS dataframe
artis <- dbGetQuery(con, "SELECT * FROM all_estimates_snet") %>%
  select(-record_id)

# consumption dataframe
consumption <- dbGetQuery(con, "SELECT * FROM complete_consumption") %>%
  select(-record_id)

# Production dataframe
prod <- dbGetQuery(con, "SELECT * FROM production") %>%
  select(-record_id)

# Product metadata dataframe
product <- dbGetQuery(con, "SELECT * FROM products") %>%
  select(-record_id)

# Species metadata
sciname_metadata <- dbGetQuery(con, "SELECT * FROM sciname") %>%
  select(-record_id)

# Improved taxa resolution metadata
sciname_hs_modified <- dbGetQuery(con, "SELECT * FROM code_max_resolved_taxa") %>%
  select(-record_id)

# Close database connection
dbDisconnect(con)
rm(list = c("con"))

artis <- artis %>%
  # Add variables for whether source/importer/exporter imposed sanctions following annexation of Crimea
    mutate(sanctions_towards_russia = case_when(
      importer_iso3c %in% c("USA", "AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", "FRA", "DEU", "GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX", "MLT", "NLD", "POL", "PRT", "ROM", "SVK", "SVN", "ESP", "SWE", "GBR", "JPN", "ALB", "ISL", "AUS", "MNT", "UKR", "CAN", "NOR", "CHE", "MDA")  ~ "Yes",
      TRUE ~ "No"
    )) %>%
      mutate(exporter_sanctions_towards_russia = case_when(
      exporter_iso3c %in% c("USA", "AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", "FRA", "DEU", "GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX", "MLT", "NLD", "POL", "PRT", "ROM", "SVK", "SVN", "ESP", "SWE", "GBR", "JPN", "ALB", "ISL", "AUS", "MNT", "UKR", "CAN", "NOR", "CHE", "MDA")  ~ "Yes",
      TRUE ~ "No")) %>% 
  left_join(sciname_metadata %>% 
              select(sciname, common_name), 
            by = "sciname") %>% 
  mutate(source_sanctions_towards_russia = case_when(
      source_country_iso3c %in% c("USA", "AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", "FRA", "DEU", "GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX", "MLT", "NLD", "POL", "PRT", "ROM", "SVK", "SVN", "ESP", "SWE", "GBR", "JPN", "ALB", "ISL", "AUS", "MNT", "UKR", "CAN", "NOR", "CHE", "MDA")  ~ "Yes",
      TRUE ~ "No")) %>% 
  # Join on common names
  left_join(sciname_metadata %>% 
              select(sciname, common_name), 
            by = "sciname")

consumption <- consumption %>%
  # Filter consumption to when Russia is involved
  filter(source_country_iso3c == "RUS"|exporter_iso3c == "RUS"|consumer_iso3c == "RUS") %>%
  # Add variables for whether source/importer/exporter imposed sanctions following annexation of Crimea
    mutate(sanctions_towards_russia = case_when(
      consumer_iso3c %in% c("USA", "AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", "FRA", "DEU", "GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX", "MLT", "NLD", 
                            "POL", "PRT", "ROM", "SVK", "SVN", "ESP", "SWE", "GBR", "JPN", "ALB", "ISL", "AUS", "MNT", "UKR", "CAN", "NOR", "CHE", "MDA")  ~ "Russia unfriendly countries",
      TRUE ~ "Russia friendly countries"
    )) %>%
      mutate(exporter_sanctions_towards_russia = case_when(
      exporter_iso3c %in% c("USA", "AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", "FRA", "DEU", "GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX", "MLT", "NLD", "POL", "PRT", "ROM", "SVK", "SVN", "ESP", "SWE", "GBR", "JPN", "ALB", "ISL", "AUS", "MNT", "UKR", "CAN", "NOR", "CHE", "MDA")  ~ "Russia unfriendly countries",
      TRUE ~ "Russia friendly countries")) %>%
    mutate(source_sanctions_towards_russia = case_when(
      source_country_iso3c %in% c("USA", "AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", "FRA", "DEU", "GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX", "MLT", "NLD", "POL", "PRT", "ROM", "SVK", "SVN", "ESP", "SWE", "GBR", "JPN", "ALB", "ISL", "AUS", "MNT", "UKR", "CAN", "NOR", "CHE", "MDA")  ~ "Russia unfriendly countries",
      TRUE ~ "Russia friendly countries")) %>% 
  # Add common names (for both sciname and sciname_hs_modified)
    left_join(sciname_metadata %>% 
              select(sciname, common_name), 
            by = "sciname") %>%
  left_join(sciname_metadata %>% 
              select(sciname, "common_name_hs_modified" = "common_name"), 
            by = c("sciname_hs_modified" = "sciname"))

```

```{r}
# Consumption of Russian products by sanctioning and non-sanctioning countries, colored by import path (direct vs indirect)
# Fill version
consumption_russian_ts_fill <- consumption %>%
  filter(consumer_iso3c != "RUS", 
         source_country_iso3c == "RUS") %>%
  mutate(path = case_when(
    exporter_iso3c == "RUS" ~ "Direct import", 
    TRUE ~ "Import from intermediate partner"
  )) %>% 
  group_by(year, path, sanctions_towards_russia) %>%
  summarise(consumption_live_t = sum(consumption_live_t)) %>%
  ggplot(aes(x = year, y = consumption_live_t/1000000, fill = path)) +
  geom_area() +
  geom_vline(aes(xintercept = 2014), linetype = 2) +
  scale_fill_manual(values = c("#D4B95f", # khaki
                               "#4FA2A2" # teal
  )) + 
  labs(x = "", y = "Consumption of Russian seafood \n(mil t, live weight)", fill = "") +
  facet_wrap(~sanctions_towards_russia) +
  theme_minimal() +
  theme(legend.position = "bottom") # FIXIT: year labels shouldn't overlap

consumption_russian_ts_fill

# Consumption of Russian products by sanctioning and non-sanctioning countries, colored by import path (direct vs indirect)
# Line version
consumption_russian_ts_line <- consumption %>%
  filter(consumer_iso3c != "RUS", 
         source_country_iso3c == "RUS") %>%
  mutate(path = case_when(
    exporter_iso3c == "RUS" ~ "Direct import", 
    TRUE ~ "Import from intermediate partner"
  )) %>% 
  group_by(year, path, sanctions_towards_russia) %>%
  summarise(consumption_live_t = sum(consumption_live_t)) %>%
  ggplot(aes(x = year, y = consumption_live_t/1000000, color = path)) +
  geom_line() +
  geom_vline(aes(xintercept = 2014), linetype = 2) +
  scale_color_manual(values = c("#D4B95f", # khaki
                               "#4FA2A2" # teal
  )) + 
  labs(x = "", y = "Consumption of Russian seafood \n(mil t, live weight)", color = "") +
  facet_wrap(~sanctions_towards_russia) +
  theme_minimal() +
  theme(legend.position = "bottom") # FIXIT: year labels shouldn't overlap

consumption_russian_ts_line

# Change in consumption stats
# Change in total consumption of Russian products - imposed sanctions
consumption_change <- consumption %>%
  filter(sanctions_towards_russia == "Russia unfriendly countries", 
         source_country_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

q_pre <- consumption_change %>% filter(year %in% 2013:2014) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)
q_post <- consumption_change %>% filter(year %in% 2015:2016) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)

100*(q_post - q_pre)/q_pre

# Change in consumption of Russian products via direct import - imposed sanctions
consumption_change <- consumption %>%
  filter(sanctions_towards_russia == "Russia unfriendly countries", 
         exporter_iso3c == "RUS",
         source_country_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

q_pre <- consumption_change %>% filter(year %in% 2013:2014) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)
q_post <- consumption_change %>% filter(year %in% 2015:2016) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)

100*(q_post - q_pre)/q_pre

# Change in consumption of Russian products via indirect import - imposed sanctions
consumption_change <- consumption %>%
  filter(sanctions_towards_russia == "Russia unfriendly countries", 
         exporter_iso3c != "RUS",
         source_country_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

q_pre <- consumption_change %>% filter(year %in% 2013:2014) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)
q_post <- consumption_change %>% filter(year %in% 2015:2016) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)

paste("Change in consumption of Russian products via indirect import - imposed sanctions")
100*(q_post - q_pre)/q_pre

# Change in total consumption of Russian products  - didn't impose sanctions
consumption_change <- consumption %>%
  filter(sanctions_towards_russia == "Russia friendly countries", 
         source_country_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

q_pre <- consumption_change %>% filter(year %in% 2013:2014) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)
q_post <- consumption_change %>% filter(year %in% 2015:2016) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)

paste("Change in total consumption of Russian products  - didn't impose sanctions")
100*(q_post - q_pre)/q_pre

```

```{r}
# Russian consumption of seafood products from sanctioning and non-sanctioning countries, colored by import path (direct vs indirect)

# Fill version
consumption_by_russia_ts_fill <- consumption %>%
  filter(consumer_iso3c == "RUS", 
         source_country_iso3c != "RUS") %>%
  mutate(path = case_when(
    source_country_iso3c == exporter_iso3c ~ "Direct import", 
    TRUE ~ "Import from intermediate partner"
  )) %>% 
  group_by(year, path, source_sanctions_towards_russia) %>%
  summarise(consumption_live_t = sum(consumption_live_t)) %>%
  ggplot(aes(x = year, y = consumption_live_t/1000000, fill = path)) +
  geom_area() +
  geom_vline(aes(xintercept = 2014), linetype = 2) +
  scale_fill_manual(values = c("#D4B95f", # khaki
                               "#4FA2A2" # teal
  )) + 
  labs(x = "", y = "Russian consumption of foreign seafood \n(mil t, live weight)", fill = "") +
  facet_wrap(~source_sanctions_towards_russia) +
  theme_minimal() +
  theme(legend.position = "bottom")

consumption_by_russia_ts_fill

# Russian consumption of seafood products from sanctioning and non-sanctioning countries, colored by import path (direct vs indirect)
# Line version
consumption_by_russia_ts_line <- consumption %>%
  filter(consumer_iso3c == "RUS", 
         source_country_iso3c != "RUS") %>%
  mutate(path = case_when(
    source_country_iso3c == exporter_iso3c ~ "Direct import", 
    TRUE ~ "Import from intermediate partner"
  )) %>% 
  group_by(year, path, source_sanctions_towards_russia) %>%
  summarise(consumption_live_t = sum(consumption_live_t)) %>%
  ggplot(aes(x = year, y = consumption_live_t/1000000, color = path)) +
  geom_line() +
  geom_vline(aes(xintercept = 2014), linetype = 2) +
  scale_color_manual(values = c("#D4B95f", # khaki
                               "#4FA2A2" # teal
  )) + 
  labs(x = "", y = "Russian consumption of foreign seafood \n(mil t, live weight)", color = "") +
  facet_wrap(~source_sanctions_towards_russia) +
  theme_minimal() +
  theme(legend.position = "bottom") # FIXIT: year labels shouldn't overlap

consumption_by_russia_ts_line

# Change in Russia's consumption from countries imposing sanctions
consumption_change <- consumption %>%
  filter(source_sanctions_towards_russia == "Russia unfriendly countries", 
         consumer_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

q_pre <- consumption_change %>% filter(year %in% 2013:2014) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)
q_post <- consumption_change %>% filter(year %in% 2015:2016) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)

paste("Change in Russia's consumption from countries imposing sanctions")
100*(q_post - q_pre)/q_pre

# Change in Russia's consumption from countries imposing sanctions - direct imports
consumption_change <- consumption %>%
  filter(source_sanctions_towards_russia == "Russia unfriendly countries", 
         source_country_iso3c == exporter_iso3c,
         consumer_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

q_pre <- consumption_change %>% filter(year %in% 2013:2014) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)
q_post <- consumption_change %>% filter(year %in% 2015:2016) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)

paste("Change in Russia's consumption from countries imposing sanctions - direct imports")
100*(q_post - q_pre)/q_pre

```
```{r, fig.height=6}
# Join plots
ggarrange(consumption_russian_ts_fill, consumption_by_russia_ts_fill, 
          common.legend = TRUE, ncol = 1)

ggarrange(consumption_russian_ts_line, consumption_by_russia_ts_line, 
          common.legend = TRUE, ncol = 1)
```

```{r}
# Sankey diagram of flows of Russian products into the US
rus_us_sankey <- consumption %>%
  filter(consumer_iso3c == "USA", source_country_iso3c == "RUS", year %in% 2015:2019) %>%
  rename("importer_iso3c" = "consumer_iso3c",
         "live_weight_t" = "consumption_live_t") %>%
  plot_sankey(prop_flow_cutoff = 0.01)

ggexport(rus_us_sankey, filename = "rus_us_sankey.png", width = 800, height = 400)

# Top species US consumes sourced from Russia
species_rank <- consumption %>%
  filter(consumer_iso3c == "USA", source_country_iso3c == "RUS", year %in% 2015:2019) %>%
  mutate(path = case_when(
    exporter_iso3c == "RUS" ~ "Direct import", 
    TRUE ~ "Import from intermediate partner"
  )) %>% 
  rename("importer_iso3c" = "consumer_iso3c",
         "live_weight_t" = "consumption_live_t") %>%
  plot_bar(bar_group = "common_name_hs_modified", fill_type = "path") + 
  theme(legend.position = "bottom")

ggexport(species_rank, filename = "US_species_rank.png")

# Stats for paper
# Percent by exporter
consumption %>%
  filter(consumer_iso3c == "USA", source_country_iso3c == "RUS", year %in% 2015:2019) %>% 
  group_by(exporter_iso3c) %>% 
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(percent = 100*consumption_live_t/sum(consumption_live_t)) %>%
  arrange(desc(percent))

# Percent by species
consumption %>%
  filter(consumer_iso3c == "USA", source_country_iso3c == "RUS", year %in% 2015:2019) %>% 
  group_by(common_name_hs_modified) %>% 
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(percent = 100*consumption_live_t/sum(consumption_live_t)) %>%
  arrange(desc(percent))

# FIXIT:Pull out stats from these data frames to print the one for the paper

```

```{r}
rus_to_chn <- artis %>%
  filter(source_country_iso3c == "RUS", exporter_iso3c == "RUS", importer_iso3c == "CHN") %>%
  group_by(year, sciname) %>%
  summarise(russia_live_t = sum(live_weight_t, na.rm = TRUE))

notrus_to_chn <- artis %>%
  filter(source_country_iso3c != "RUS", exporter_iso3c != "RUS", importer_iso3c == "CHN") %>%
  group_by(year, sciname) %>%
  summarise(notrussia_live_t = sum(live_weight_t, na.rm = TRUE))

chn_to_usa <- artis %>%
  filter(exporter_iso3c == "CHN", importer_iso3c == "USA") %>%
  group_by(year, sciname) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
  left_join(rus_to_chn, by = c("year", "sciname")) %>%
  left_join(notrus_to_chn, by = c("year", "sciname")) %>%
  replace_na(list(russia_live_t = 0, notrussia_live_t = 0)) %>% 
  # Calculate the minimum by year and species that had to come from Russia
  mutate(min_russia = case_when(
    # If non-Russian product exceeds the volume sent to the US, 
    # then min that had to be Russian is 0
    (notrussia_live_t >= live_weight_t) ~ 0,
    # If non-Russian product cannot explain all, then fill up remaining with Russian product
    (live_weight_t - notrussia_live_t) >= russia_live_t ~ russia_live_t,
    # Otherwise fill up remaining with Russian product
    TRUE ~ (live_weight_t - notrussia_live_t)
  )) %>% 
  # Calculate the max that could have come from Russia
  mutate(max_russia = case_when(
    # If Russian product can explain full amount, max is total trade flow
    russia_live_t >= live_weight_t ~ live_weight_t, 
    # If Russian product cannont explain all, fill up with Russian product to get max
    russia_live_t < live_weight_t ~ russia_live_t
  ))

# FIXIT: This can be streamlined to not repeat code
p1 <- artis %>%
  filter(source_country_iso3c == "RUS", exporter_iso3c == "CHN", importer_iso3c == "USA") %>%
  group_by(year, sciname) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>% 
  filter(sciname == "gadus chalcogrammus") %>%
  left_join(chn_to_usa %>%
              select(year, sciname, min_russia, max_russia)) %>% 
  ggplot(aes(x = year)) +
  geom_ribbon(aes(ymin = min_russia, ymax = max_russia), fill = "#86ADA7", alpha = 0.5) +
  geom_line(aes(y = live_weight_t)) +
  labs(x = "", y = "tonnes, live weight", title = "US imports of Russian Alaska pollock via China") +
  theme_bw()

p2 <- artis %>%
  filter(source_country_iso3c == "RUS", exporter_iso3c == "CHN", importer_iso3c == "USA") %>%
  group_by(year, sciname) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>% 
  filter(sciname == "oncorhynchus gorbuscha") %>%
  left_join(chn_to_usa %>%
              select(year, sciname, min_russia, max_russia)) %>% 
  ggplot(aes(x = year)) +
  geom_ribbon(aes(ymin = min_russia, ymax = max_russia), fill = "#86ADA7", alpha = 0.5) +
  geom_line(aes(y = live_weight_t)) +
  labs(x = "", y = "tonnes, live weight", title = "US imports of Russian pink salmon via China") +
  theme_bw()

p3 <- artis %>%
  filter(source_country_iso3c == "RUS", exporter_iso3c == "CHN", importer_iso3c == "USA",
         year > 1996) %>%
  group_by(year, sciname) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>% 
  filter(sciname == "gadus morhua") %>%
  left_join(chn_to_usa %>%
              select(year, sciname, min_russia, max_russia)) %>% 
  ggplot(aes(x = year)) +
  geom_ribbon(aes(ymin = min_russia, ymax = max_russia), fill = "#86ADA7", alpha = 0.5) +
  geom_line(aes(y = live_weight_t)) +
  labs(x = "", y = "tonnes, live weight", title = "US imports of Russian Atlantic cod via China") +
  theme_bw()

p4 <- artis %>%
  filter(source_country_iso3c == "RUS", exporter_iso3c == "CHN", importer_iso3c == "USA",
         year > 1996) %>%
  group_by(year, sciname) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>% 
  filter(sciname == "melanogrammus aeglefinus") %>%
  left_join(chn_to_usa %>%
              select(year, sciname, min_russia, max_russia)) %>% 
  ggplot(aes(x = year)) +
  geom_ribbon(aes(ymin = min_russia, ymax = max_russia), fill = "#86ADA7", alpha = 0.5) +
  geom_line(aes(y = live_weight_t)) +
  labs(x = "", y = "tonnes, live weight", title = "US imports of Russian haddock via China") +
  theme_bw()

p5 <- artis %>%
  filter(source_country_iso3c == "RUS", exporter_iso3c == "CHN", importer_iso3c == "USA",
         year > 1996) %>%
  group_by(year, sciname) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>% 
  filter(sciname == "gadus macrocephalus") %>%
  left_join(chn_to_usa %>%
              select(year, sciname, min_russia, max_russia)) %>% 
  ggplot(aes(x = year)) +
  geom_ribbon(aes(ymin = min_russia, ymax = max_russia), fill = "#86ADA7", alpha = 0.5) +
  geom_line(aes(y = live_weight_t)) +
  labs(x = "", y = "tonnes, live weight", title = "US imports of Russian Pacific cod via China") +
  theme_bw()

p_joined <- ggarrange(p1, p2, p3, p4, ncol = 1)

ggexport(p_joined, filename = "US_species_import_range.png")

# Stats
chn_to_usa %>% 
  filter(year == 2020, 
         sciname %in% c("gadus chalcogrammus", "gadus morhua", "oncorhynchus gorbuscha", "gadus macrocephalus"))

consumption %>%
  filter(consumer_iso3c == "USA", source_country_iso3c == "RUS", year %in% 2015:2019) %>%
  group_by(sciname_hs_modified) %>% 
  summarise(consumption_live_t = sum(consumption_live_t)) %>% 
  arrange(desc(consumption_live_t)) %>%
  ungroup() %>%
  mutate(percent = 100*consumption_live_t/sum(consumption_live_t)) %>%
  mutate(cum_percent = cumsum(percent))

```

```{r}
# Explore spike in China's apparent consumption of Russian products
tmp <- consumption %>% 
  filter(source_country_iso3c == "RUS", consumer_iso3c != "RUS") %>%
  group_by(year, consumer_iso3c) %>% 
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

consumption %>% 
  filter(source_country_iso3c == "RUS", consumer_iso3c == "CHN") %>%
  group_by(year) %>% 
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = consumption_live_t)) +
  geom_line() + 
  theme_bw()

tmp <- consumption %>% 
  filter(source_country_iso3c == "RUS", consumer_iso3c == "CHN") %>%
  group_by(year, sciname_hs_modified) %>% 
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

prod %>% 
  filter(iso3c == "RUS", sciname == "gadus chalcogrammus") %>%
  group_by(year) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ggplot(aes(x = year, y = live_weight_t)) +
  geom_line() +
  theme_bw()

artis %>% 
  filter(exporter_iso3c == "RUS", importer_iso3c == "CHN") %>%
  plot_ts(artis_var = "hs6", plot.type = "filled", prop_flow_cutoff = 0.07)

artis %>% 
  filter(source_country_iso3c == "RUS", exporter_iso3c == "CHN") %>%
  group_by(year) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
  ggplot(aes(year, live_weight_t)) +
  geom_line() +
  theme_bw()

artis %>% 
  filter(source_country_iso3c == "RUS", exporter_iso3c == "CHN") %>%
  plot_ts(artis_var = "importer_iso3c", plot.type = "filled")

artis %>% 
  filter(hs6 %in% c(30379, 30380, 303429, 30499, 30510, 30520, 
                    30530, 30549, 30569, 51191, 1400410, 140420,
                    160419, 160430), exporter_iso3c == "CHN") %>%
  plot_ts(artis_var = "hs6", plot.type = "filled")

```