### 2024-03-14 ms-russia-figures.Rmd
figure edits

---
title: "Russian sanctions figures"
output: html_document
date: "2024-02-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To do: 
-

```{r}
library(tidyverse)
library(countrycode)
library(exploreARTIS)
library(DBI)
library(janitor)
library(broom)
library(dbscan)
library(GGally)
library(ggrepel)
library(ggpubr)
```

```{r}
habitat_source_colors <- c("#1B708F", # blue = marine capture
                          "#86ADA7", # light blue = freshwater capture
                          "#F37542", # medium orange = marine aquaculture
                          "#F7AF75", # light orange = inland aquaculture
                          "grey50" # unknown
                          )

region_colors <- c(
  "#741A32", # maroon
  "#B34232", # burnt orange
  "#D38F35", #  orange
  "#D4B95f", # khaki
  "#4FA2A2", # teal
  "#114F59" # dark teal
)
```

```{r}
# Initial database pulls

# Database connection
con <- dbConnect(RPostgres::Postgres(),
                 dbname=Sys.getenv("DB_NAME"),
                 host="localhost",
                 port="5432",
                 user=Sys.getenv("DB_USERNAME"),
                 password=Sys.getenv("DB_PASSWORD"))

# Check that connection is established by checking which tables are present
dbListTables(con)

# ARTIS dataframe
artis <- dbGetQuery(con, "SELECT * FROM snet") %>%
  select(-record_id)

# consumption dataframe
consumption <- dbGetQuery(con, "SELECT * FROM complete_consumption") %>%
  select(-record_id)

# Production dataframe
prod <- dbGetQuery(con, "SELECT * FROM production") %>%
  select(-record_id)

# Product metadata dataframe
product <- dbGetQuery(con, "SELECT * FROM products") %>%
  select(-record_id)

# Species metadata
sciname_metadata <- dbGetQuery(con, "SELECT * FROM sciname") %>%
  select(-record_id)

# Improved taxa resolution metadata
sciname_hs_modified <- dbGetQuery(con, "SELECT * FROM code_max_resolved_taxa") %>%
  select(-record_id)

# Close database connection
dbDisconnect(con)
rm(list = c("con"))

artis <- artis %>%
  # Add variables for whether source/importer/exporter imposed sanctions following annexation of Crimea
    mutate(sanctions_towards_russia = case_when(
      importer_iso3c %in% c("USA", "AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", "FRA", "DEU", "GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX", "MLT", "NLD", "POL", "PRT", "ROM", "SVK", "SVN", "ESP", "SWE", "GBR", "JPN", "ALB", "ISL", "AUS", "MNT", "UKR", "CAN", "NOR", "CHE", "MDA")  ~ "Yes",
      TRUE ~ "No"
    )) %>%
      mutate(exporter_sanctions_towards_russia = case_when(
      exporter_iso3c %in% c("USA", "AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", "FRA", "DEU", "GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX", "MLT", "NLD", "POL", "PRT", "ROM", "SVK", "SVN", "ESP", "SWE", "GBR", "JPN", "ALB", "ISL", "AUS", "MNT", "UKR", "CAN", "NOR", "CHE", "MDA")  ~ "Yes",
      TRUE ~ "No")) %>% 
  left_join(sciname_metadata %>% 
              select(sciname, common_name), 
            by = "sciname") %>% 
  mutate(source_sanctions_towards_russia = case_when(
      source_country_iso3c %in% c("USA", "AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", "FRA", "DEU", "GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX", "MLT", "NLD", "POL", "PRT", "ROM", "SVK", "SVN", "ESP", "SWE", "GBR", "JPN", "ALB", "ISL", "AUS", "MNT", "UKR", "CAN", "NOR", "CHE", "MDA")  ~ "Yes",
      TRUE ~ "No")) %>% 
  # Join on common names
  left_join(sciname_metadata %>% 
              select(sciname, common_name), 
            by = "sciname")

consumption <- consumption %>%
  # Filter consumption to when Russia is involved
  filter(source_country_iso3c == "RUS"|exporter_iso3c == "RUS"|consumer_iso3c == "RUS") %>%
  # Add variables for whether source/importer/exporter imposed sanctions following annexation of Crimea
    mutate(sanctions_towards_russia = case_when(
      consumer_iso3c %in% c("USA", "AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", "FRA", "DEU", "GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX", "MLT", 
                            "NLD", # FIXIT: Need to find out what is happening with the Netherlands and others directly importing
                            "POL", "PRT", "ROM", "SVK", "SVN", "ESP", "SWE", "GBR", "JPN", "ALB", "ISL", "AUS", "MNT", "UKR", "CAN", "NOR", "CHE", "MDA")  ~ "Imposed sanctions",
      TRUE ~ "Did not impose sanctions"
    )) %>%
      mutate(exporter_sanctions_towards_russia = case_when(
      exporter_iso3c %in% c("USA", "AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", "FRA", "DEU", "GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX", "MLT", "NLD", "POL", "PRT", "ROM", "SVK", "SVN", "ESP", "SWE", "GBR", "JPN", "ALB", "ISL", "AUS", "MNT", "UKR", "CAN", "NOR", "CHE", "MDA")  ~ "Imposed sanctions",
      TRUE ~ "Did not impose sanctions")) %>%
    mutate(source_sanctions_towards_russia = case_when(
      source_country_iso3c %in% c("USA", "AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", "FRA", "DEU", "GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX", "MLT", "NLD", "POL", "PRT", "ROM", "SVK", "SVN", "ESP", "SWE", "GBR", "JPN", "ALB", "ISL", "AUS", "MNT", "UKR", "CAN", "NOR", "CHE", "MDA")  ~ "Imposed sanctions",
      TRUE ~ "Did not impose sanctions")) %>% 
  # Add common names (for both sciname and sciname_hs_modified)
    left_join(sciname_metadata %>% 
              select(sciname, common_name), 
            by = "sciname") %>%
  left_join(sciname_metadata %>% 
              select(sciname, "common_name_hs_modified" = "common_name"), 
            by = c("sciname_hs_modified" = "sciname"))

```

```{r}
# Consumption of Russian products by sanctioning and non-sanctioning countries, colored by import path (direct vs indirect)
# Fill version
consumption_of_rus_p <- consumption %>%
  filter(consumer_iso3c != "RUS", 
         source_country_iso3c == "RUS") %>%
  mutate(path = case_when(
    exporter_iso3c == "RUS" ~ "Direct import", 
    TRUE ~ "Import from intermediate partner"
  )) %>% 
  group_by(year, path, sanctions_towards_russia) %>%
  summarise(consumption_live_t = sum(consumption_live_t)) %>%
  ggplot(aes(x = year, y = consumption_live_t/1000000, fill = path)) +
  geom_area() +
  geom_vline(aes(xintercept = 2014), linetype = 2) +
  scale_fill_manual(values = c("#D4B95f", # khaki
                               "#4FA2A2" # teal
  )) + 
  labs(x = "", y = "Consumption of Russian seafood \n(mil t, live weight)", fill = "") +
  facet_wrap(~sanctions_towards_russia) +
  theme_minimal() +
  theme(
    axis.title.y = element_text(size = 9),
    legend.position = "top",
    panel.spacing = unit(5, "mm")
  )

consumption_of_rus_p

# Consumption of Russian products by sanctioning and non-sanctioning countries, colored by import path (direct vs indirect)
# Line version
consumption %>%
  filter(consumer_iso3c != "RUS", 
         source_country_iso3c == "RUS") %>%
  mutate(path = case_when(
    exporter_iso3c == "RUS" ~ "Direct import", 
    TRUE ~ "Import from intermediate partner"
  )) %>% 
  group_by(year, path, sanctions_towards_russia) %>%
  summarise(consumption_live_t = sum(consumption_live_t)) %>%
  ggplot(aes(x = year, y = consumption_live_t/1000000, color = path)) +
  geom_line() +
  geom_vline(aes(xintercept = 2014), linetype = 2) +
  scale_color_manual(values = c("#D4B95f", # khaki
                               "#4FA2A2" # teal
  )) + 
  labs(x = "", y = "Consumption of Russian seafood \n(mil t, live weight)", color = "") +
  facet_wrap(~sanctions_towards_russia) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    panel.spacing = unit(5, "mm")
  )

# Change in consumption stats
# Change in total consumption of Russian products - imposed sanctions
consumption_change <- consumption %>%
  filter(sanctions_towards_russia == "Imposed sanctions", 
         source_country_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

q_pre <- consumption_change %>% filter(year %in% 2013:2014) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)
q_post <- consumption_change %>% filter(year %in% 2015:2016) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)

100*(q_post - q_pre)/q_pre

# Change in consumption of Russian products via direct import - imposed sanctions
consumption_change <- consumption %>%
  filter(sanctions_towards_russia == "Imposed sanctions", 
         exporter_iso3c == "RUS",
         source_country_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

q_pre <- consumption_change %>% filter(year %in% 2013:2014) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)
q_post <- consumption_change %>% filter(year %in% 2015:2016) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)

100*(q_post - q_pre)/q_pre

# Change in consumption of Russian products via indirect import - imposed sanctions
consumption_change <- consumption %>%
  filter(sanctions_towards_russia == "Imposed sanctions", 
         exporter_iso3c != "RUS",
         source_country_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

q_pre <- consumption_change %>% filter(year %in% 2013:2014) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)
q_post <- consumption_change %>% filter(year %in% 2015:2016) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)

paste("Change in consumption of Russian products via indirect import - imposed sanctions")
100*(q_post - q_pre)/q_pre

# Change in total consumption of Russian products  - didn't impose sanctions
consumption_change <- consumption %>%
  filter(sanctions_towards_russia == "Did not impose sanctions", 
         source_country_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

q_pre <- consumption_change %>% filter(year %in% 2013:2014) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)
q_post <- consumption_change %>% filter(year %in% 2015:2016) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)

paste("Change in total consumption of Russian products  - didn't impose sanctions")
100*(q_post - q_pre)/q_pre

```

```{r}
# Russian consumption of seafood products from sanctioning and non-sanctioning countries, colored by import path (direct vs indirect)
# Fill version
rus_foreign_consumption_p <- consumption %>%
  filter(consumer_iso3c == "RUS", 
         source_country_iso3c != "RUS") %>%
  mutate(path = case_when(
    source_country_iso3c == exporter_iso3c ~ "Direct import", 
    TRUE ~ "Import from intermediate partner"
  )) %>% 
  group_by(year, path, source_sanctions_towards_russia) %>%
  summarise(consumption_live_t = sum(consumption_live_t)) %>%
  ggplot(aes(x = year, y = consumption_live_t/1000000, fill = path)) +
  geom_area() +
  geom_vline(aes(xintercept = 2014), linetype = 2) +
  scale_fill_manual(values = c("#D4B95f", # khaki
                               "#4FA2A2" # teal
  )) + 
  labs(x = "", y = "Russian consumption of foreign seafood \n(mil t, live weight)", fill = "") +
  facet_wrap(~source_sanctions_towards_russia) +
  theme_minimal() +
  theme(
    axis.title.y = element_text(size = 9),
    legend.position = "top",
    panel.spacing = unit(5, "mm")
  )

rus_foreign_consumption_p

# Russian consumption of seafood products from sanctioning and non-sanctioning countries, colored by import path (direct vs indirect)
# Line version
consumption %>%
  filter(consumer_iso3c == "RUS", 
         source_country_iso3c != "RUS") %>%
  mutate(path = case_when(
    source_country_iso3c == exporter_iso3c ~ "Direct import", 
    TRUE ~ "Import from intermediate partner"
  )) %>% 
  group_by(year, path, source_sanctions_towards_russia) %>%
  summarise(consumption_live_t = sum(consumption_live_t)) %>%
  ggplot(aes(x = year, y = consumption_live_t/1000000, color = path)) +
  geom_line() +
  geom_vline(aes(xintercept = 2014), linetype = 2) +
  scale_color_manual(values = c("#D4B95f", # khaki
                               "#4FA2A2" # teal
  )) + 
  labs(x = "", y = "Russian consumption of foreign seafood \n(mil t, live weight)", color = "") +
  facet_wrap(~source_sanctions_towards_russia) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    panel.spacing = unit(5, "mm")
  )

# Change in Russia's consumption from countries imposing sanctions
consumption_change <- consumption %>%
  filter(source_sanctions_towards_russia == "Imposed sanctions", 
         consumer_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

q_pre <- consumption_change %>% filter(year %in% 2013:2014) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)
q_post <- consumption_change %>% filter(year %in% 2015:2016) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)

paste("Change in Russia's consumption from countries imposing sanctions")
100*(q_post - q_pre)/q_pre

# Change in Russia's consumption from countries imposing sanctions - direct imports
consumption_change <- consumption %>%
  filter(source_sanctions_towards_russia == "Imposed sanctions", 
         source_country_iso3c == exporter_iso3c,
         consumer_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

q_pre <- consumption_change %>% filter(year %in% 2013:2014) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)
q_post <- consumption_change %>% filter(year %in% 2015:2016) %>% summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% pull(consumption_live_t)

paste("Change in Russia's consumption from countries imposing sanctions - direct imports")
100*(q_post - q_pre)/q_pre

```

```{r fig.height = 6, width = 8}
# Format figure 1 together
ggarrange(
  consumption_of_rus_p,
  rus_foreign_consumption_p,
  nrow = 2,
  common.legend = TRUE
)

ggsave("fig1.png", device = "png", height = 5, width = 6.5, units = "in")
```

```{r}
# Sankey diagram of flows of Russian products into the US
rus_us_sankey <- consumption %>%
  filter(consumer_iso3c == "USA", source_country_iso3c == "RUS", year %in% 2015:2019) %>%
  rename("importer_iso3c" = "consumer_iso3c",
         "live_weight_t" = "consumption_live_t") %>%
  plot_sankey(prop_flow_cutoff = 0.01)

ggexport(rus_us_sankey, filename = "rus_us_sankey.png", width = 800, height = 400)

# Top species US consumes sourced from Russia
species_rank <- consumption %>%
  filter(consumer_iso3c == "USA", source_country_iso3c == "RUS", year %in% 2015:2019) %>%
  mutate(path = case_when(
    exporter_iso3c == "RUS" ~ "Direct import", 
    TRUE ~ "Import from\nintermediate partner"
  )) %>% 
  rename("importer_iso3c" = "consumer_iso3c",
         "live_weight_t" = "consumption_live_t") %>%
  mutate(common_name_hs_modified = case_when(
    common_name_hs_modified == "alaska pollock(=walleye poll.)" ~ "Alaska Pollock",
    common_name_hs_modified == "atlantic cod" ~ "Atlantic Cod",
    common_name_hs_modified == "pink(=humpback) salmon" ~ "Pink Salmon",
    common_name_hs_modified == "haddock" ~ "Haddock",
    common_name_hs_modified == "pacific cod" ~ "Pacific Cod",
    TRUE ~ common_name_hs_modified
  )) %>%
  mutate(common_name_hs_modified = gsub(" ", "\n", common_name_hs_modified)) %>%
  plot_bar(bar_group = "common_name_hs_modified", fill_type = "path", top_n = 5) + 
  theme(
    legend.position = "bottom",
    axis.text.y = element_text(angle = 90),

  ) +
  labs(y = "", x = "Consumption (live weight tonnes)", fill = "")

ggexport(species_rank, filename = "US_species_rank.png")

# Stats for paper
# Percent by exporter
consumption %>%
  filter(consumer_iso3c == "USA", source_country_iso3c == "RUS", year %in% 2015:2019) %>% 
  group_by(exporter_iso3c) %>% 
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(percent = 100*consumption_live_t/sum(consumption_live_t)) %>%
  arrange(desc(percent))

# Percent by species
consumption %>%
  filter(consumer_iso3c == "USA", source_country_iso3c == "RUS", year %in% 2015:2019) %>% 
  group_by(common_name_hs_modified) %>% 
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(percent = 100*consumption_live_t/sum(consumption_live_t)) %>%
  arrange(desc(percent))

# FIXIT:Pull out stats from these data frames to print the one for the paper

```

```{r}
rus_to_chn <- artis %>%
  # Isolating Chinese imports that were originally caught in Russia
  filter(source_country_iso3c == "RUS", importer_iso3c == "CHN") %>%
  group_by(year, sciname) %>%
  summarise(russia_live_t = sum(live_weight_t, na.rm = TRUE))

notrus_to_chn <- artis %>%
  # Isolating Chinese imports that were NOT originally caught in Russia
  filter(source_country_iso3c != "RUS", importer_iso3c == "CHN") %>%
  group_by(year, sciname) %>%
  summarise(notrussia_live_t = sum(live_weight_t, na.rm = TRUE))

chn_to_usa <- artis %>%
  filter(exporter_iso3c == "CHN", importer_iso3c == "USA") %>%
  group_by(year, sciname) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
  left_join(rus_to_chn, by = c("year", "sciname")) %>%
  left_join(notrus_to_chn, by = c("year", "sciname")) %>%
  replace_na(list(russia_live_t = 0, notrussia_live_t = 0)) %>% 
  # Calculate the minimum by year and species that had to come from Russia
  mutate(min_russia = case_when(
    # If non-Russian product exceeds the volume sent to the US, 
    # then min that had to be Russian is 0
    (notrussia_live_t >= live_weight_t) ~ 0,
    # If non-Russian product cannot explain all, then fill up remaining with Russian product
    (live_weight_t - notrussia_live_t) >= russia_live_t ~ russia_live_t,
    # Otherwise fill up remaining with Russian product
    TRUE ~ (live_weight_t - notrussia_live_t)
  )) %>% 
  # Calculate the max that could have come from Russia
  mutate(max_russia = case_when(
    # If Russian product can explain full amount, max is total trade flow
    russia_live_t >= live_weight_t ~ live_weight_t, 
    # If Russian product cannont explain all, fill up with Russian product to get max
    russia_live_t < live_weight_t ~ russia_live_t
  ))

# Stats
chn_to_usa %>% 
  filter(year == 2020, 
         sciname %in% c("gadus chalcogrammus", "gadus morhua", "oncorhynchus gorbuscha", "gadus macrocephalus"))

consumption %>%
  filter(consumer_iso3c == "USA", source_country_iso3c == "RUS", year %in% 2015:2019) %>%
  group_by(sciname_hs_modified) %>% 
  summarise(consumption_live_t = sum(consumption_live_t)) %>% 
  arrange(desc(consumption_live_t)) %>%
  ungroup() %>%
  mutate(percent = 100*consumption_live_t/sum(consumption_live_t)) %>%
  mutate(cum_percent = cumsum(percent))

```

```{r}
rus_chn_usa <- artis %>%
  filter(source_country_iso3c == "RUS", exporter_iso3c == "CHN", importer_iso3c == "USA") %>%
  group_by(year, sciname) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(chn_to_usa %>%
            select(year, sciname, min_russia, max_russia)) %>%
  mutate(live_weight_t = live_weight_t / 1e3,
         min_russia = min_russia / 1e3,
         max_russia = max_russia / 1e3)

p1 <- rus_chn_usa %>% 
  filter(sciname == "gadus chalcogrammus",
         year > 1996) %>%
  ggplot(aes(x = year)) +
  geom_ribbon(aes(ymin = min_russia, ymax = max_russia), fill = "#86ADA7", alpha = 0.5) +
  geom_line(aes(y = live_weight_t)) +
  xlim(1997, 2020) +
  scale_y_continuous(position = "right") +
  labs(x = "", y = "tonnes, live weight") +#, title = "US imports of Russian Alaska pollock via China") +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    plot.margin = margin(l = 2, b = -2, unit = "mm")
  )

p2 <- rus_chn_usa %>%
  filter(year > 1996,
         sciname == "gadus morhua") %>%
  left_join(chn_to_usa %>%
              select(year, sciname, min_russia, max_russia)) %>% 
  ggplot(aes(x = year)) +
  geom_ribbon(aes(ymin = min_russia, ymax = max_russia), fill = "#86ADA7", alpha = 0.5) +
  geom_line(aes(y = live_weight_t)) +
  scale_y_continuous(position = "right") +
  xlim(1997, 2020) +
  labs(x = "", y = "tonnes, live weight") +#, title = "US imports of Russian Atlantic cod via China") +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    plot.margin = margin(r = 2, l = 2, b = -2, unit = "mm")
  )

p3 <- rus_chn_usa %>%
  filter(sciname == "oncorhynchus gorbuscha",
         year > 1996) %>%
  ggplot(aes(x = year)) +
  geom_ribbon(aes(ymin = min_russia, ymax = max_russia), fill = "#86ADA7", alpha = 0.5) +
  geom_line(aes(y = live_weight_t)) +
  scale_y_continuous(position = "right") +
  xlim(1997, 2020) +
  labs(x = "", y = "tonnes, live weight") +#, title = "US imports of Russian pink salmon via China") +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    plot.margin = margin(r = 2, l = 2, b = -2, unit = "mm")
  )

p4 <- rus_chn_usa %>%
  filter(sciname == "melanogrammus aeglefinus",
         year > 1996) %>%
  ggplot(aes(x = year)) +
  geom_ribbon(aes(ymin = min_russia, ymax = max_russia), fill = "#86ADA7", alpha = 0.5) +
  geom_line(aes(y = live_weight_t)) +
  scale_y_continuous(position = "right") +
  xlim(1997, 2020) +
  labs(x = "", y = "tonnes, live weight") +#, title = "US imports of Russian haddock via China") +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    plot.margin = margin(l = 2, b = -2, r = 1.5, unit = "mm")
  )

p5 <- rus_chn_usa %>%
  filter(sciname == "gadus macrocephalus",
         year > 1996) %>%
  ggplot(aes(x = year)) +
  geom_ribbon(aes(ymin = min_russia, ymax = max_russia), fill = "#86ADA7", alpha = 0.5) +
  geom_line(aes(y = live_weight_t)) +
  scale_y_continuous(position = "right") +
  xlim(1997, 2020) +
  labs(x = "", y = "tonnes, live weight") +#, title = "US imports of Russian Pacific cod via China") +
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    plot.margin = margin(t = -1, r = 2, l = 2, b = -2, unit = "mm")
  )

p_joined <- ggarrange(p1, p2, p3, p4, p5, ncol = 1)

ggexport(p_joined, filename = "US_species_import_range.png")

# Top species US consumes sourced from Russia

species_rank_df <- consumption %>%
  filter(consumer_iso3c == "USA", source_country_iso3c == "RUS", year %in% 2015:2019) %>%
  mutate(path = case_when(
    exporter_iso3c == "RUS" ~ "Direct\nimport", 
    TRUE ~ "Import from\nintermediate partner"
  )) %>% 
  rename("importer_iso3c" = "consumer_iso3c",
         "live_weight_t" = "consumption_live_t") %>%
  mutate(common_name_hs_modified = case_when(
    common_name_hs_modified == "alaska pollock(=walleye poll.)" ~ "Alaska Pollock",
    common_name_hs_modified == "atlantic cod" ~ "Atlantic Cod",
    common_name_hs_modified == "pink(=humpback) salmon" ~ "Pink Salmon",
    common_name_hs_modified == "haddock" ~ "Haddock",
    common_name_hs_modified == "pacific cod" ~ "Pacific Cod",
    TRUE ~ common_name_hs_modified
  )) %>%
  mutate(common_name_hs_modified = gsub(" ", "\n", common_name_hs_modified)) %>%
  mutate(live_weight_t = live_weight_t / length(2015:2019)) %>%
  mutate(live_weight_t = live_weight_t / 1e3)

species_rank_1 <- species_rank_df %>%
  filter(common_name_hs_modified == "Alaska\nPollock") %>%
  mutate(dom_source == "") %>%
  plot_bar(bar_group = "common_name_hs_modified", fill_type = "path", top_n = 5) +
  xlim(0, 90) +
  theme(
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.text.y = element_text(angle = 90, hjust = 0.5),
    plot.margin = margin(b = -2, unit = "mm")
  ) +
  labs(y = "", x = "", fill = "")

species_rank_2 <- species_rank_df %>%
  filter(common_name_hs_modified == "Atlantic\nCod") %>%
  plot_bar(bar_group = "common_name_hs_modified", fill_type = "path", top_n = 5) + 
  xlim(0, 90) +
  theme(
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.text.y = element_text(angle = 90, hjust = 0.5),
    plot.margin = margin(b = -2, unit = "mm")
  ) +
  labs(y = "", x = "", fill = "")

species_rank_3 <- species_rank_df %>%
  filter(common_name_hs_modified == "Pink\nSalmon") %>%
  plot_bar(bar_group = "common_name_hs_modified", fill_type = "path", top_n = 5) + 
  xlim(0, 90) +
  theme(
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.text.y = element_text(angle = 90, hjust = 0.5),
    plot.margin = margin(b = -2, unit = "mm")
  ) +
  labs(y = "", x = "", fill = "")

species_rank_4 <- species_rank_df %>%
  filter(common_name_hs_modified == "Haddock") %>%
  mutate(common_name_hs_modified = paste("\n", common_name_hs_modified, sep = "")) %>%
  plot_bar(bar_group = "common_name_hs_modified", fill_type = "path", top_n = 5) + 
  xlim(0, 90) +
  theme(
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.text.y = element_text(angle = 90, hjust = 0.5),
    plot.margin = margin(b = -2, unit = "mm")
  ) +
  labs(y = "", x = "", fill = "")

species_rank_5 <- species_rank_df %>%
  filter(common_name_hs_modified == "Pacific\nCod") %>%
  plot_bar(bar_group = "common_name_hs_modified", fill_type = "path", top_n = 5) + 
  xlim(0, 90) +
  theme(
    legend.position = "bottom",
    axis.text.y = element_text(angle = 90, hjust = 0.5),
    axis.title.x = element_text(size = 7),
    legend.key.size = unit(1, units = "line"),
    plot.margin = margin(t = -2, b = -1, unit = "mm")
  ) +
  labs(y = "", x = "Avg. Consumption (1000 t) (2015-2019)", fill = "")

# Format Figure 2 together
fig2 <- ggarrange(
  species_rank_1, p1,
  species_rank_2, p2,
  species_rank_3, p3,
  species_rank_4, p4,
  species_rank_5, p5,
  ncol = 2, nrow = 5,
  widths = c(1, 2),
  heights = c(1, 1, 1, 1, 1),
  common.legend = TRUE,
  legend = "bottom"
  )

fig2

fig2 <- annotate_figure(
  fig2,
  right = text_grob("US Imports of Russian Species (1000 t, live weight)", rot = 270)
  )

fig2 <- ggarrange(
  rus_us_sankey,
  fig2,
  nrow = 2,
  heights = c(1, 2)
)

fig2

ggsave("fig2.png", device = "png", width = 6.5, height = 6)
```

```{r}
# Explore spike in China's apparent consumption of Russian products
tmp <- consumption %>% 
  filter(source_country_iso3c == "RUS", consumer_iso3c != "RUS") %>%
  group_by(year, consumer_iso3c) %>% 
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

consumption %>% 
  filter(source_country_iso3c == "RUS", consumer_iso3c == "CHN") %>%
  group_by(year) %>% 
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = consumption_live_t)) +
  geom_line() + 
  theme_bw()

tmp <- consumption %>% 
  filter(source_country_iso3c == "RUS", consumer_iso3c == "CHN") %>%
  group_by(year, sciname_hs_modified) %>% 
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE))

prod %>% 
  filter(iso3c == "RUS", sciname == "gadus chalcogrammus") %>%
  group_by(year) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ggplot(aes(x = year, y = live_weight_t)) +
  geom_line() +
  theme_bw()

artis %>% 
  filter(exporter_iso3c == "RUS", importer_iso3c == "CHN") %>%
  plot_ts(artis_var = "hs6", plot.type = "filled", prop_flow_cutoff = 0.07)

artis %>% 
  filter(source_country_iso3c == "RUS", exporter_iso3c == "CHN") %>%
  group_by(year) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
  ggplot(aes(year, live_weight_t)) +
  geom_line() +
  theme_bw()

artis %>% 
  filter(source_country_iso3c == "RUS", exporter_iso3c == "CHN") %>%
  plot_ts(artis_var = "importer_iso3c", plot.type = "filled")

artis %>% 
  filter(hs6 %in% c(30379, 30380, 303429, 30499, 30510, 30520, 
                    30530, 30549, 30569, 51191, 1400410, 140420,
                    160419, 160430), exporter_iso3c == "CHN") %>%
  plot_ts(artis_var = "hs6", plot.type = "filled")

```