### 2024-06-11 ms-russia-results.Rmd

---
title: "ms_russia_results"
author: ""
date: "2024-04-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(include = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

library(tidyverse)
library(reactable)
library(reactablefmtr)

datadir <- "data"
```

```{r load_data}
consumption <- read.csv(file.path(datadir, "case_study_consumption.csv")) %>%
  mutate(sciname_hs_modified = case_when(
    is.na(sciname_hs_modified) ~ sciname,
    TRUE ~ sciname_hs_modified
  ))

consumption_by_rus <- read.csv(file.path(datadir, "fig1_consumption_by_russia.csv"))

rus_chn_usa <- read.csv(file.path(datadir, "rus_chn_usa.csv"))

us_consumption <- read.csv(file.path(datadir, "case_study_usa_consumption.csv")) %>%
  mutate(sciname_hs_modified = case_when(
    is.na(sciname_hs_modified) ~ sciname,
    TRUE ~ sciname_hs_modified
  ))
```

```{r rus_imports_change}
# Calculating Russia's consumption from countries imposing sanctions across years
consumption_change <- consumption %>%
  filter(source_sanctions_towards_russia == "Russia unfriendly countries",
         consumer_iso3c == "RUS") %>%
  group_by(year) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>%
  ungroup()

# Russia's consumption (2013-2014) from countries imposing sanctions
q_pre <- consumption_change %>%
  filter(year %in% 2013:2014) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>%
  ungroup() %>%
  pull(consumption_live_t)

# Russia's consumption (2013-2014) from countries imposing sanctions
q_post <- consumption_change %>%
  filter(year %in% 2015:2016) %>%
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>%
  ungroup %>%
  pull(consumption_live_t)

# Percent change in Russian consumption from countries imposing sanctions
percent_change <- 100 * ((q_post - q_pre) / q_pre)
```

Russia's consumption, in 2013-2014, from countries imposing sanctions was `r signif(q_pre / 1e6, digits = 3)` million tonnes. Russia's consumption, in 2015-2016, from countries imposing sanctions was `r signif(q_post / 1e5, digits = 3)` hundred thousand tonnes. This represented a percent change of `r signif(percent_change, digits = 3)`%.

```{r top_intermediate_partners, echo = TRUE}
intermediate_partners <- consumption %>%
  filter(consumer_iso3c == "USA",
         source_country_iso3c == "RUS",
         year %in% 2015:2019) %>% 
  group_by(exporter_iso3c) %>% 
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(percent = 100 * consumption_live_t / sum(consumption_live_t)) %>%
  arrange(desc(percent))

top_intermediates <- intermediate_partners %>%
  slice_max(order_by = percent, n = 5)

colnames(top_intermediates) <- c("Intermediate Exporter", "Consumption (live weight tonnes)", "Percent (%)")
```

### Top Intermediate Exporters for Russian sourced products to the US (2015-2019)
`r reactable(top_intermediates)`

```{r}
top_intermediate_partner <- intermediate_partners %>%
  slice_max(order_by = percent, n = 1) %>%
  pull(exporter_iso3c)

top_intermediate_partner_percent <- intermediate_partners %>%
  slice_max(order_by = percent, n = 1) %>%
  pull(percent)

top_intermediate_partner_species <- consumption %>%
  filter(consumer_iso3c == "USA", exporter_iso3c == top_intermediate_partner, source_country_iso3c == "RUS", year %in% 2015:2019) %>% 
  group_by(common_name_hs_modified) %>% 
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(percent = 100*consumption_live_t/sum(consumption_live_t)) %>%
  arrange(desc(percent)) %>%
  slice_max(order_by = percent, n = 3)
```

### Top Species Exported from RUS -> `r top_intermediate_partner` -> USA
`r reactable(top_intermediate_partner_species)`

We show that in recent years (2015-2019), `r signif(top_intermediate_partner_percent, digits = 3)` of Russian products consumed in the US moved through `r top_intermediate_partner`, with the majority being `r top_intermediate_partner_species[1,]$common_name_hs_modified`, `r top_intermediate_partner_species[2,]$common_name_hs_modified`, and `r top_intermediate_partner_species[3,]$common_name_hs_modified`.

```{r unfriendly_countries_consumption}
unfriendly_rus_consumption <- consumption %>%
  filter(consumer_iso3c != "RUS" &
           sanctions_towards_russia == "Russia unfriendly countries" &
           ((year %in% 2013:2014) | (year %in% 2015:2016))) %>%
  mutate(period = case_when(
    year %in% 2013:2014 ~ "before_sanctions",
    TRUE ~ "after_sanctions"
  )) %>%
  group_by(sanctions_towards_russia, period) %>%
  summarize(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>%
  ungroup()

unfriendly_consumption_pre <- unfriendly_rus_consumption %>%
  filter(period == "before_sanctions") %>%
  pull(consumption_live_t)

unfriendly_consumption_post <- unfriendly_rus_consumption %>%
  filter(period == "after_sanctions") %>%
  pull(consumption_live_t)

unfriendly_consumption_percent_change = 100 * ((unfriendly_consumption_post - unfriendly_consumption_pre) / unfriendly_consumption_pre)
```

Analyzing consumption of Russian-caught seafood following the 2014 round of sanctions, we find that countries that imposed sanctions (“unfriendly countries”) initially moderately reduced total consumption of Russian seafood (`r signif(unfriendly_consumption_percent_change, digits = 3)`%) in 2015–2016 compared to 2013-2014.

```{r friendly_countries_consumption}
# Look at 2014 drop in consumption of Russian products by friendly countries
friendly_rus_consumption <- consumption %>%
  filter(consumer_iso3c != "RUS" & 
           sanctions_towards_russia == "Russia friendly countries" &
           ((year %in% 2012:2013) | (year %in% 2014))) %>%
  mutate(period = case_when(
    year %in% 2012:2013 ~ "before_sanctions",
    TRUE ~ "after_sanctions"
  )) %>%
  group_by(sanctions_towards_russia, period) %>%
  summarize(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)/length(unique(year))) %>%
  ungroup()

friendly_consumption_pre <- friendly_rus_consumption %>%
  filter(period == "before_sanctions") %>%
  pull(consumption_live_t)

friendly_consumption_post <- friendly_rus_consumption %>%
  filter(period == "after_sanctions") %>%
  pull(consumption_live_t)

friendly_consumption_percent_change = 100 * ((friendly_consumption_post - friendly_consumption_pre) / friendly_consumption_pre)

# Friendly country consumption of Russian products following 2014
friendly_consumption_2020 <- consumption %>%
  filter(consumer_iso3c != "RUS" & 
           sanctions_towards_russia == "Russia friendly countries" &
           year == 2020) %>%
  group_by(year) %>%
  summarize(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>%
  pull(consumption_live_t)

friendly_consumption_percent_change_2020 = 100 * ((friendly_consumption_2020 - friendly_consumption_pre) / friendly_consumption_pre)
```

Countries which did not impose sanctions (“friendly countries”) decreased consumption of Russian products (`r signif(friendly_consumption_percent_change, digits = 3)`% in 2014 compared to 2012–2013), but their consumption quickly rebounded, exceeding `r floor(friendly_consumption_percent_change_2020)`% of the pre-2014 value in 2020.

```{r consumption_by_russia}

consumption_by_rus <- consumption %>%
  filter(consumer_iso3c == "RUS", 
         source_country_iso3c != "RUS",
         ((year %in% 2013:2014) | (year %in% 2015:2016))) %>%
  mutate(period = case_when(
    year %in% 2013:2014 ~ "before_sanctions",
    TRUE ~ "after_sanctions"
  )) %>%
  group_by(source_sanctions_towards_russia, period) %>%
  summarise(consumption_live_t = sum(consumption_live_t)) %>%
  ungroup()

rus_of_unfriendly <- consumption_by_rus %>%
  filter(source_sanctions_towards_russia == "Russia unfriendly countries")

rus_of_unfriendly_pre <- rus_of_unfriendly %>%
  filter(period == "before_sanctions") %>%
  pull(consumption_live_t)

rus_of_unfriendly_post <- rus_of_unfriendly %>%
  filter(period == "after_sanctions") %>%
  pull(consumption_live_t)

rus_of_unfriendly_percent_change <- 100 * ((rus_of_unfriendly_post - rus_of_unfriendly_pre) / rus_of_unfriendly_pre)
```

### Consumption By Russia
`r reactable(consumption_by_rus)`

Russia reduced consumption of seafood from unfriendly countries by `r signif(rus_of_unfriendly_percent_change, digits = 3)`% between 2015-2016 compared to 2013-2014.

```{r}
rus_to_us_path <- consumption %>%
  filter(consumer_iso3c == "USA" & source_country_iso3c == "RUS" &
           (year >= 2015 & year <= 2019)) %>%
  group_by(source_country_iso3c, exporter_iso3c, consumer_iso3c) %>%
  summarize(consumption_live_t = sum(consumption_live_t)) %>%
  ungroup() %>%
  mutate(path = case_when(
    exporter_iso3c == "RUS" ~ "Direct",
    TRUE ~ "Indirect"
  )) %>%
  group_by(source_country_iso3c, consumer_iso3c, path) %>%
  summarize(consumption_live_t = sum(consumption_live_t)) %>%
  ungroup() %>%
  group_by(consumer_iso3c, source_country_iso3c) %>%
  mutate(percent = 100 * (consumption_live_t/sum(consumption_live_t))) %>%
  ungroup()

rus_to_us_indirect_percent <- rus_to_us_path %>%
  filter(path == "Indirect") %>%
  pull(percent)
```

### US Direct vs Indirect consumption of Russian Products (2015-2019)

`r reactable(rus_to_us_path)`

We find that for 2015–2019, `r signif(rus_to_us_indirect_percent, digits = 3)`% of Russian-harvested seafood consumed in the US was processed in and imported from a country other than Russia, with `r signif(top_intermediate_partner_percent, digits = 3)`% moving through `r top_intermediate_partner`

```{r}

rus_to_us_species <- consumption %>%
  filter(consumer_iso3c == "USA", source_country_iso3c == "RUS", year %in% 2015:2019) %>% 
  group_by(sciname_hs_modified, common_name_hs_modified) %>% 
  summarise(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>% 
  ungroup() %>%
  mutate(percent = 100 * consumption_live_t / sum(consumption_live_t)) %>%
  arrange(desc(percent))

top_rus_to_us_species <- rus_to_us_species %>%
  slice_max(order_by = percent, n = 5)

top_rus_to_us_species_percent <- sum(top_rus_to_us_species$percent)
```

The top five species were `r top_rus_to_us_species[1,]$common_name_hs_modified` (`r top_rus_to_us_species[1,]$sciname_hs_modified`), `r top_rus_to_us_species[2,]$common_name_hs_modified` (`r top_rus_to_us_species[2,]$sciname_hs_modified`), `r top_rus_to_us_species[3,]$common_name_hs_modified` (`r top_rus_to_us_species[3,]$sciname_hs_modified`),  `r top_rus_to_us_species[4,]$common_name_hs_modified` (`r top_rus_to_us_species[4,]$sciname_hs_modified`), and `r top_rus_to_us_species[5,]$common_name_hs_modified` (`r top_rus_to_us_species[5,]$sciname_hs_modified`), which collectively made up `r signif(top_rus_to_us_species_percent, digits = 3)`% of US consumption of Russian products.

```{r consumption_minimums_2020}
# Min value from Russia
alaska_pollock_min <- rus_chn_usa %>%
  filter(year == 2020 &
           sciname == "gadus chalcogrammus") %>%
  pull(min_russia)

pink_salmon_min <- rus_chn_usa %>%
  filter(year == 2020 &
           sciname == "oncorhynchus gorbuscha") %>%
  pull(min_russia)

# Max value from Russia
alaska_pollock_max <- rus_chn_usa %>%
  filter(year == 2020 &
           sciname == "gadus chalcogrammus") %>%
  pull(max_russia)

pink_salmon_max <- rus_chn_usa %>%
  filter(year == 2020 &
           sciname == "oncorhynchus gorbuscha") %>%
  pull(max_russia)

us_total_pollock_consumption <- us_consumption %>%
  filter(sciname == "gadus chalcogrammus") %>%
  # Min/max flow values are in 1000 tonnes
  summarise(consumption_live_t = sum(consumption_live_t) / 1e3) %>%
  pull(consumption_live_t)

us_imported_pollock_consumption <- us_consumption %>%
  filter(sciname == "gadus chalcogrammus",
         !is.na(exporter_iso3c)) %>%
  # Min/max flow values are in 1000 tonnes
  summarise(consumption_live_t = sum(consumption_live_t) / 1e3) %>%
  pull(consumption_live_t)

us_imported_salmon_consumption <- us_consumption %>%
  filter(sciname == "oncorhynchus gorbuscha",
         !is.na(exporter_iso3c)) %>%
  # Min/max flow values are in 1000 tonnes
  summarise(consumption_live_t = sum(consumption_live_t) / 1e3) %>%
  pull(consumption_live_t)

us_russian_pollock_consumption <- us_consumption %>%
  filter(sciname == "gadus chalcogrammus",
         !is.na(exporter_iso3c),
         source_country_iso3c == "RUS") %>%
  # Min/max flow values are in 1000 tonnes
  summarise(consumption_live_t = sum(consumption_live_t) / 1e3) %>%
  pull(consumption_live_t)

us_russian_salmon_consumption <- us_consumption %>%
  filter(sciname == "oncorhynchus gorbuscha",
         !is.na(exporter_iso3c),
         source_country_iso3c == "RUS") %>%
  # Min/max flow values are in 1000 tonnes
  summarise(consumption_live_t = sum(consumption_live_t) / 1e3) %>%
  pull(consumption_live_t)

us_consumption %>%
  filter(consumer_iso3c == "USA", 
         sciname_hs_modified %in% c("oncorhynchus gorbuscha", "gadus chalcogrammus")) %>%
  mutate(consumption_source = case_when(
    (source_country_iso3c == consumer_iso3c & is.na(exporter_iso3c)) ~ "domestic",
    (source_country_iso3c == consumer_iso3c & !is.na(exporter_iso3c)) ~ "domestic, foreign processed",
    (source_country_iso3c != consumer_iso3c) ~ "foreign"
  )) %>%
  group_by(sciname_hs_modified, consumption_source) %>%
  summarise(consumption_live_t = sum(consumption_live_t)) %>%
  group_by(sciname_hs_modified) %>%
  mutate(percent = 100*consumption_live_t/sum(consumption_live_t))

# tmp <- us_consumption %>%
#   filter(consumer_iso3c == "USA", 
#          sciname_hs_modified %in% c("oncorhynchus gorbuscha", "gadus chalcogrammus")) %>%
#   group_by(sciname_hs_modified, source_country_iso3c) %>%
#   summarise(consumption_live_t = sum(consumption_live_t)) %>%
#   group_by(sciname_hs_modified) %>%
#   mutate(percent = 100*consumption_live_t/sum(consumption_live_t))

us_consumption %>%
  filter(consumer_iso3c == "USA", 
         sciname_hs_modified == "oncorhynchus gorbuscha") %>%
  mutate(exporter_iso3c = case_when(
    is.na(exporter_iso3c) ~ source_country_iso3c,
    TRUE ~ exporter_iso3c
  )) %>%
  plot_sankey(cols = c("source_country_iso3c", 
                       "exporter_iso3c", 
                       "consumer_iso3c"),
              value = "consumption_live_t")

```

While US imports of Atlantic cod from China could technically be explained without Russian-sourced products, we estimate that the US imported at least `r signif(alaska_pollock_min, digits = 3)` thousand and `r signif(pink_salmon_min, digits = 3)` thousand tonnes of Russian-sourced Alaska pollock and pink salmon, respectively, via China in 2020.

Based on these ranges, we estimate `r signif(100*alaska_pollock_min/us_imported_pollock_consumption, digits = 3)`-`r min(signif(100*alaska_pollock_max/us_imported_pollock_consumption, digits = 3), 100)`% of imported Alaska pollock and `r signif(100*pink_salmon_min/us_imported_salmon_consumption, digits = 3)`-`r min(signif(100*pink_salmon_max/us_imported_salmon_consumption, digits = 3), 100)`% of imported pink salmon is Russian-harvested.